---
title: "additional_analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(xtable)

full_raw <- read.table("tiamat1-full.out", header=T)
extra_raw <- read.table("tiamat1-extra.out", header=T)
coefs_raw <- read.table("tiamat1-coefs2.out", header=T)
singular <- read.table("singular_results.out")
old_raw <- read.table("tiamat1-oldresults.out", header=T)
medium_raw <- read.table("tiamat1-medium.out", header=T)
selected_raw <- read.table("tiamat1-selected.out", header=T)
```

```{r}
full <- full_raw %>%
  group_by(instance, algorithm, reducer) %>%
  summarize_all(mean, na.rm=T) %>%
  select(-rep)

extra <- extra_raw %>%
  group_by(instance, algorithm, reducer) %>%
  summarize_all(mean, na.rm=T) %>%
  select(-rep)

coefs <- coefs_raw %>%
  group_by(instance, algorithm, reducer) %>%
  summarize_all(mean, na.rm=T) %>%
  select(-rep)

old <- old_raw %>%
  group_by(instance, algorithm, reducer) %>%
  summarize_all(mean, na.rm=T) %>%
  select(-rep)

medium <- medium_raw %>%
  group_by(instance, algorithm, reducer) %>%
  summarize_all(mean, na.rm=T) %>%
  select(-rep)

selected <- selected_raw %>%
  group_by(instance, algorithm, reducer) %>%
  summarize_all(mean, na.rm=T) %>%
  select(-rep)

full
extra
coefs
selected
```

Plotting coefficient growth to have some ideas.

```{r}
ggplot(coefs, aes(x=instance, y=totalcoefs, fill=algorithm)) + geom_col(position='dodge') + coord_flip() + scale_y_log10()
```

Checking out if there's anything interesting to show about the Singular results:

```{r}
colnames(singular) <- c("instance", "time", "polynomials", "monomials", "degree")
singular
```

Answer: not really, but I can show some results if that's important.

Defining geometric means (used in the tables in the paper):

```{r}
#We drop zeroes in the geometric mean, it is important in some cases with overhead 0
gmean <- function(x, na.rm=FALSE) {
  exp(mean(log(x[which(x != 0)]), na.rm=na.rm))
}
```

Remaking tables for the paper (only includes the initial 30 instances - I think I won't use this one anymore, and only use the big table below):

```{r}
caption <- "Geometric means of the experiments over all instances (ignoring timeouts)."
t <- old %>% 
  group_by(algorithm, reducer) %>%
  summarize(
    timeout=sum(is.na(time)),
    time=gmean(time, na.rm=TRUE),
    overhead=gmean(dynamic, na.rm=TRUE),
    polynomials=gmean(polynomials, na.rm=TRUE),
    monomials=gmean(monomials, na.rm=TRUE),
    degree=gmean(degree, na.rm=TRUE),
    sreductions=gmean(sreductions, na.rm=TRUE)
  ) %>%
  select(
    algorithm, reducer, time, overhead, polynomials, 
    monomials, degree, sreductions,timeout
  ) %>%
    rename(
      "Algorithm" = algorithm,
      "Reducer" = reducer,
      "$t$" = time,
      "$O$" = overhead,
      "$|G|$" = polynomials,
      "$|\\Supp(G)|$" = monomials,
      "$\\deg$" = degree,
      "sred" = sreductions
    )
  print(xtable(t,caption=caption), 
        include.rownames=F, 
        sanitize.text.function=function(x){x},
        NA.string = "NA")
```

Generating a table to check coefficient growth.

```{r}
caption <- "Geometric means of the experiments over all instances (ignoring timeouts) over $\\mathbb{Q}$, including total coefficient bitsize."
t <- coefs %>% 
  group_by(algorithm) %>%
  summarize(
    timeout=sum(is.na(time)),
    time=gmean(time, na.rm=TRUE),
    totalcoefs=gmean(totalcoefs, na.rm=TRUE)
  ) %>%
  select(
    algorithm, time, totalcoefs, timeout
  ) %>%
    rename(
      "Algorithm" = algorithm,
      "$t$" = time,
      "coef" = totalcoefs
    )
  print(xtable(t,caption=caption),
        include.rownames=F, 
        sanitize.text.function=function(x){x},
        NA.string = "NA")
```

Current results for coefficient tests are a bit weird. Why are bases using grevlex smaller?
In short, because of the case where Static times out (cyclicnh7). In reality, results for polynomials / monomials / degree are all very similar to the
modular case.

butcher8: all algorithms timed out
cyclicn7: all algorithms timed out
cyclicnh*: all dynamic algorithms win easily here, static timed out
noon7: everything BUT static timed out
katsuran7: cp timed out here

```{r}
coefs %>% filter(instance == "cyclicnh5")
coefs %>% filter(instance == "cyclicnh6")
coefs %>% filter(instance == "cyclicnh7")
coefs
```

Checking results / table that could be generated for the "full" set. 84 new instances, this sounds useful. What about complete results, using the old
instances too?

```{r}
unique(full$instance)
useless <- c("cyclicn2", "cyclicn3", "cyclicn4-simple", "cyclicn5-permut", "cyclicn5-simple", "cyclicnh2", "cyclicnh3",
             "katsuran2", "katsuran3", "katsuran4-simple", "katsuran6-simple", "katsuranh2", "katsuranh3", "econ2", "econ3",
             "econh2", "econh3")
full <- full %>% filter(!(instance %in% useless))
length(unique(full$instance))

all <- rbind(full, old)
all <- rbind(all, selected)
all
```

This table looks nice for the paper, including all new results (up until now, at least). It includes 114 ideals!

```{r}
length(unique(all$instance))
```

```{r}
caption <- "Geometric means of the experiments over all instances (ignoring timeouts)."
t <- all %>% 
  group_by(algorithm, reducer) %>%
  summarize(
    timeout=sum(is.na(time)),
    time=gmean(time, na.rm=TRUE),
    overhead=gmean(dynamic, na.rm=TRUE),
    polynomials=gmean(polynomials, na.rm=TRUE),
    monomials=gmean(monomials, na.rm=TRUE),
    degree=gmean(degree, na.rm=TRUE),
    sreductions=gmean(sreductions, na.rm=TRUE)
  ) %>%
  select(
    algorithm, reducer, time, overhead, polynomials, 
    monomials, degree, sreductions,timeout
  ) %>%
    rename(
      "Algorithm" = algorithm,
      "Reducer" = reducer,
      "$t$" = time,
      "$O$" = overhead,
      "$|G|$" = polynomials,
      "$|\\Supp(G)|$" = monomials,
      "$\\deg$" = degree,
      "sred" = sreductions
    )
  print(xtable(t,caption=caption), 
        include.rownames=F, 
        sanitize.text.function=function(x){x},
        NA.string = "NA")
```

Compute the dynamic overhead in the cases where the dynamic algorithms were unable
to find anything better than grevlex.
Result: the overhead is large for small instances, but when we take only those that took longer than 1s to run, the situation changes, and the overhead
tends to be lower than 10%.

```{r}
all %>% 
  filter(reducer == "classical", time > 1) %>%
  select(instance, algorithm, time, dynamic, polynomials) %>% 
  pivot_wider(names_from = c(algorithm), values_from = c(time, polynomials, dynamic)) %>%
  filter(polynomials_perturbation == polynomials_static) %>%
  mutate(overhead = dynamic_perturbation / time_perturbation) %>%
  select(instance, overhead)

all %>% 
  filter(reducer == "F4", time > 1) %>%
  select(instance, algorithm, time, dynamic, polynomials) %>% 
  pivot_wider(names_from = c(algorithm), values_from = c(time, polynomials, dynamic)) %>%
  filter(polynomials_perturbation == polynomials_static) %>%
  mutate(overhead = dynamic_perturbation / time_perturbation) %>%
  select(instance, overhead)

all %>% 
  filter(reducer == "classical", time > 1) %>%
  select(instance, algorithm, time, dynamic, polynomials) %>% 
  pivot_wider(names_from = c(algorithm), values_from = c(time, polynomials, dynamic)) %>%
  filter(polynomials_random == polynomials_static) %>%
  mutate(overhead = dynamic_random / time_random) %>%
  select(instance, overhead)

all %>% 
  filter(reducer == "F4", time > 1) %>%
  select(instance, algorithm, time, dynamic, polynomials) %>% 
  pivot_wider(names_from = c(algorithm), values_from = c(time, polynomials, dynamic)) %>%
  filter(polynomials_random == polynomials_static) %>%
  mutate(overhead = dynamic_random / time_random) %>%
  select(instance, overhead)
```

I find the new scatterplot with even more points than before more confusing than helpful, to be honest. I won't include it in the new version.

```{r}
p <- ggplot(data=all, aes(x=polynomials, y=time, color=algorithm)) +
  geom_point() +
  scale_y_log10() +
  scale_x_log10() +
  xlab("log(Polynomials)") +
  ylab("log(Time)") +
  guides(color=guide_legend("Algorithm")) +
  theme(legend.position=c(0.87, 0.23))
#ggsave(paste0("./img/scatterpolytime.pdf"), height=4.5)
plot(p)
```

Verifying queue recomputation ratio:

```{r}
queue <- all %>% filter(!(is.na(time))) %>% mutate(ratio = queue / time) %>% select(instance, algorithm, reducer, ratio)
queue
queue %>% group_by(algorithm, reducer) %>% summarize(ratio = mean(ratio))
```
